---
title: "Competititon III. Correspondence analysis"
author: "Roma Siugzdaite"
date: "2025-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Competition III.

HairEyeColor dataset. Distribution of hair and eye color and sex in 592 statistics students.

```{r }

data(HairEyeColor)
HairEyeColor
hair_eye = as.data.frame(HairEyeColor) # just keep your data in a data frame

```

## Make a contingency table

```{r}
#  contingency table for hair and eye color
library("FactoMineR")
library("factoextra")
library(ISLR)
library(tidyverse)
library(Rfast)
library(MASS)

hair_eye.table <- prop.table(hair_eye.table)
addmargins(hair_eye.table)

```

## Plot a mosaicplot

Prepare frequencies to plot 
```{r}
library("FactoMineR")
library("factoextra")
hair_eye.table <- xtabs(Freq ~ Eye + Hair, data=hair_eye)
mosaicplot(hair_eye.table, color = (1:6), main = "Eye vs Hair")

```

### Chi square test

```{r}

chisq.test(hair_eye.table)

```


## Run Correspondence Analysis

```{r}
CA(hair_eye.table, ncp = 5, graph = TRUE)

res.ca <- CA(hair_eye.table, graph = FALSE)

eig.val <- get_eigenvalue(res.ca)
eig.val


```



## Plot CA in 2 dimensions

```{r}
fviz_ca_biplot(res.ca, repel = TRUE)
```


Interpretation???

####################### 

## NEW DATA Children

####################### 

The data used here is a contingency table that summarizes the answers given by different categories of people to the following question : according to you, what are the reasons that can make hesitate a woman or a couple to have children? A data frame with 18 rows and 8 columns. Rows represent the different reasons mentioned, columns represent the different categories (education, age) people belong to.

```{r}

library(FactoMineR)
data(children)
children
children <- as.data.frame(children) # just keep your data in a data frame

res.ca <- CA(children, row.sup = 15:18, col.sup = 6:8)
summary(res.ca)


## Ellipses for 5 the active elements
ellipseCA(res.ca,ellipse="col",col.col.ell=c(rep("blue",5),rep("transparent",3)), invisible=c("row.sup","col.sup"))

```

Interpretation?

# Third dataset: Titanic

Question: How we can look at Survival depending on the Class of the tickets? 

Question: How Survival depends on Gender?

```{r}
Titanic
Titanic <- as.data.frame(Titanic) # just keep your data 
```


