---
title: "MultivariateAnalysis"
output:
  html_document: default
  pdf_document: default
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Beer data

cost is representing the PCA analysis of bear data - analysis of factors 


```{r}

```



```{r }

Beer <- read.csv("~/Desktop/r-project/Multivariate Analysis/Beer.txt")

head(Beer, 5)

summary(Beer)

#checking missing data
colSums(is.na(Beer))
missings <- colSums(is.na(Beer))
summary(missings)

#histogram of bear and cost
hist(Beer$cost)
shapiro.test(Beer$cost)

plot(Beer$cost, Beer$reputat)

```
There is no correlation because the p value is too small which is against the null hypothesis that cost values are normally distributed.

## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}

library(RcmdrMisc)
library(psych)

rcorr.adjust(Beer)


Beer2 <- na.omit(Beer)


r <- rcorr.adjust(Beer2, c("pearson"))

corPlot(r$R$r, numbers=TRUE)
```

## KMO Test
Kaiser provided the following values for interpreting the results:

* 0.00 to 0.49 unacceptable
* 0.50 to 0.59 miserable
* 0.60 to 0.69 mediocre
* 0.70 to 0.79 middling
* 0.80 to 0.89 meritorious
* 0.90 to 1.00 marvelous


```{r}

library(psych)
KMO(Beer2)

```
```{r}

BeerClean <- Beer2[, KMO(Beer2)$MSAi > 0.515] # Get rid of all variables with MSA < 0.50, to get rid of social economic status because it is not related to the beer itself.

KMO(BeerClean)$MSAi
round( KMO(BeerClean)$MSA, 2 )

```

```{r}
library(psych)
cortest.bartlett(BeerClean)

```
if p<0.05, then we reject Null hypothesis that there are no common factors in my data. and we accet Alternative Hypothesis that there are common factors.

```{r}

scree(BeerClean, pc=FALSE) 

scree(BeerClean, pc=TRUE)
```
```{r}
ev <- eigen(cor(BeerClean)) # get eigenvalues
ev$values
```

```{r}
fa.parallel(BeerClean, fa="fa")
```
```{r}
Nfacs <- 2 # This is for four factors. You can change this as needed.

fit <- factanal(BeerClean, Nfacs, rotation="promax")


print(fit, digits=2, cutoff=0.3, sort=TRUE)
```

```{r}
load <- fit$loadings[,1:2]
plot(load,type="p") # set up plot
text(load,labels=names(BeerClean),cex=.7)

```
```{r}
library(psych)

loads <- fit$loadings

fa.diagram(loads)
```


```{r}

dim(fit$loadings)

round(fit$loadings[1:6,], 2)
```


```{r}
FactorLoadings <- round(fit$loadings[1:6,], 2)

write.csv(FactorLoadings, file="FacLoads.csv")
```


