---
title: "soda"
output: html_document
date: "2025-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r cars}
Soda <- read.csv("~/Desktop/r-project/Multivariate Analysis/Soda.txt", dec=",")
head(Soda, 12)
summary(Soda)

#checking missing data 
colSums(is.na(Soda))
missings <- colSums(is.na(Soda))
summary(missings)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#histogram of soda and factors
Soda$X1.Refreshing.taste
hist(Soda$X1.Refreshing.taste)
shapiro.test(Soda$X1.Refreshing.taste)

Soda$X2.Fewer.calories
hist(Soda$X2.Fewer.calories)
shapiro.test(Soda$X2.Fewer.calories)

Soda$X3.Quenches.thirst
hist(Soda$X3.Quenches.thirst)
shapiro.test(Soda$X3.Quenches.thirst)

Soda$X4.Sweet.taste
hist(Soda$X4.Sweet.taste)
shapiro.test(Soda$X4.Sweet.taste)

Soda$X5.Energy
hist(Soda$X5.Energy)
shapiro.test(Soda$X5.Energy)

Soda$X6.Environmentally.friendly.packaging
hist(Soda$X6.Environmentally.friendly.packaging)
shapiro.test(Soda$X6.Environmentally.friendly.packaging)

Soda$X7.Minerals.and.vitamins
hist(Soda$X7.Minerals.and.vitamins)
shapiro.test(Soda$X7.Minerals.and.vitamins)

Soda$X8.Unique.flavor
hist(Soda$X8.Unique.flavor)
shapiro.test(Soda$X8.Unique.flavor)

Soda$X9.Healthy
hist(Soda$X9.Healthy)
shapiro.test(Soda$X9.Healthy)

Soda$X10.For.when.very.thirsty
hist(Soda$X10.For.when.very.thirsty)
shapiro.test(Soda$X10.For.when.very.thirsty)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
plot(Soda$X3.Quenches.thirst, Soda$X10.For.when.very.thirsty)
```
```{r}
#calculating correlations between 10 variables
library(RcmdrMisc)
library(psych)

Soda2 <- na.omit(Soda)
SodaClean<-Soda2[,3:12]

rcorr.adjust(SodaClean)





r <- rcorr.adjust(SodaClean, c("pearson"))

corPlot(r$R$r, numbers=TRUE, cex = 0.5)
```
#calculating KMO value
```{r}

library(psych)
KMO(SodaClean)

```


```{r}
SodaFinal <- SodaClean[, KMO(SodaClean)$MSAi > 0.515] # Get rid of all variables with MSA < 0.50
KMO(SodaFinal)$MSAi
round( KMO(SodaFinal)$MSA, 2 )
```

```{r}
cortest.bartlett(SodaFinal)
```

```{r}
scree(SodaFinal, pc=FALSE) 

scree(SodaFinal, pc=TRUE)
```


```{r}
ev <- eigen(cor(SodaFinal)) # get eigenvalues
ev$values
```

#FA

```{r}
fa.parallel(SodaFinal, fa="fa")
```


```{r}
Nfacs <- 3 # This is for four factors. You can change this as needed.

fit <- factanal(SodaFinal, Nfacs, rotation="promax")


print(fit, digits=2, cutoff=0.3, sort=TRUE)
```

when factor numbers are three, the model gives the optimal predictability and it indicates that the included factors are sufficient to account for the variation. 

```{r}
load <- fit$loadings[,1:2]
plot(load,type="p") # set up plot
text(load,labels=names(SodaFinal),cex=.7)
```
there are clustering of factors, soda taste-related factors tend to gather at the top left. Health and environment related factors gather around the bottom left, leaving other factors around the bottom right. 

```{r}
loads <- fit$loadings

fa.diagram(loads)
```


```{r}
dim(fit$loadings)

round(fit$loadings[1:6,], 2)
```
```{r}
FactorLoadings <- round(fit$loadings[1:6,], 2)

write.csv(FactorLoadings, file="SodaFacLoads.csv")
```

